---
title: "Exploratory Data Analysis"
author: "Michal"
date: "2024-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(dplyr)
```

## Imports
```{r}
#knitr::purl("Data_Cleaning.Rmd", output = #"cleaning_function.R")
```

```{r}
source("cleaning_function.R")
```

```{r}
uncleaned_bots = read.csv("C:/Users/Owner/Desktop/Data Science Project/Twitter-Bot-Code/archive/twitter_human_bots_dataset.csv")
```

## Data

```{r}
bots_cleaned = cleaning_function(uncleaned_bots)
head(bots_cleaned)
```


## Univariate Analysis

This section looks through the available columns and plots to introduce the data.


### Created At Date

Year
```{r}
ggplot(data = bots_cleaned, mapping = aes(year(created_at)))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Year Created Distribution")+
  xlab("Year Created")+
  ylab("Count")+
  theme_minimal()
```


Month
```{r}
ggplot(data = bots_cleaned, mapping = aes(month(created_at, label=TRUE)))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Month Created Distribution")+
  xlab("Month Created")+
  ylab("Count")+
  theme_minimal()
```


Months in 2019
```{r}
ggplot(data = bots_cleaned%>%filter(year(created_at)== 2019), mapping = aes(month(created_at, label=TRUE)))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Month in 2019 Created Distribution")+
  xlab("Month Created")+
  ylab("Count")+
  theme_minimal()
```


Weekday
```{r}
ggplot(data = bots_cleaned, mapping = aes(wday(created_at, label=TRUE)))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Weekday Created Distribution")+
  xlab("Weekday Created")+
  ylab("Count")+
  theme_minimal()
```

*Observations*

* The year created distribution is skewed right, with a peak in 2009.
* The monthly data is approximately uniform with slight spikes for March, April, and June, and a decrease towards the end of the year
* There 2019 data is only complete through April.
* There are slightly more accounts created on Wednesdays than the rest of the weekdays. The weekends have the least accounts created, but the difference is not large


### Default Profile

```{r}
ggplot(data = bots_cleaned, mapping = aes(default_profile))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Default Profile Distribution")+
  xlab("Has Default Profile")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Somewhat less than half of the accounts have default profiles


### Default Profile Image

```{r}
ggplot(data = bots_cleaned, mapping = aes(default_profile_image))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Default Profile Image Distribution")+
  xlab("Has Default Profile Image")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Very few accounts have the default profile image


### Description

Has Description
```{r}
ggplot(data = bots_cleaned, mapping = aes(str_length(description)!=0))+
  geom_histogram(stat="count",fill="darkblue", color = "black")+
  ggtitle("Accounts with Description")+
  xlab("Has a Description")+
  ylab("Count")+
  theme_minimal()
```

Description length
```{r}
ggplot(data = bots_cleaned, mapping = aes(str_length(description)))+
  geom_histogram(fill="darkblue", color = "black", binwidth = 1)+
  ggtitle("Description Length Histogram")+
  xlab("Description Length")+
  ylab("Count")+
  theme_minimal()
```

Description length excluding no description
```{r}
ggplot(data = bots_cleaned%>%filter(str_length(description)!=0), mapping = aes(str_length(description)))+
  geom_histogram(fill="blue", color = "black", binwidth = 1)+
  ggtitle("Description Length Excluding No Description")+
  xlab("Description Length")+
  ylab("Count")+
  theme_minimal()

ggplot(data = bots_cleaned%>%filter(str_length(description)!=0), mapping = aes(str_length(description)))+
  geom_boxplot(fill="blue", color = "black")+
  ggtitle("Description Length Box Plot")+
  xlab("Description Length")+
  theme(panel.background = element_rect(fill = "white"))+
  theme_minimal()
```

Number of spaces in description among accounts with description
```{r}
ggplot(data = bots_cleaned%>%filter(description!=""), mapping = aes(str_count(description,"\\s")))+
  geom_histogram(fill="blue", color = "black", binwidth = 1)+
  ggtitle("Space Count Histogram")+
  xlab("Space Count")+
  ylab("Count")+
  theme_minimal()
```

Periods in Description (accounts with description)
```{r}
ggplot(data = bots_cleaned%>%filter(description!=""), mapping = aes(str_count(description,"\\.")))+
  geom_histogram(fill="blue", color = "black", binwidth = 1)+
  ggtitle("Description Period Count")+
  xlab("Number of Periods in Description")+
  ylab("Count")+
  theme_minimal()
```

Punctuation in Description (accounts with description)
```{r}
ggplot(data = bots_cleaned%>%filter(description!=""), mapping = aes(str_count(description,"[:punct:]")))+
  geom_histogram(fill="blue", color = "black", binwidth = 1)+
  ggtitle("Description Punctuation Count")+
  xlab("Number of Punctuation in Description")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Most accounts have a description, but a significant minority do not
* There is a somewhat bimodal distribution of description length, with peaks around 25 and 160 characters
* Space count, Period count, and Punctuation count is skewed right with a peak at 0


### Geography Enabled

```{r}
ggplot(data = bots_cleaned, mapping = aes(geo_enabled))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Geography Enabled Distribution")+
  xlab("Has Geography Enabled")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* A slight majority of accounts do not have geography enabled


### id

```{r}
ggplot(data = bots_cleaned, mapping = aes(id))+
  geom_histogram(fill="blue", color = "black")+
  ggtitle("User ID Distribution")+
  xlab("User ID")+
  ylab("Count")+
  scale_x_log10()+
  theme_minimal()
```
*Observations*

* Id values are strongly skewed right


### Language


```{r}
lang_count=bots_cleaned%>%
  count(lang)%>%
  slice_max(n,n=2)
lang_count

rest=bots_cleaned%>%
  count(lang)%>%
  anti_join(lang_count)%>%
  summarise(sum(n))
rest

final=bind_rows(lang_count,rest)%>%
  mutate(lang=if_else(lang=="", "None", lang))%>%
  mutate(lang=if_else(is.na(lang), "Other", lang))%>%  
  mutate(n=if_else(is.na(n), 0, n))%>%
  mutate(`sum(n)`=if_else(is.na(`sum(n)`), 0, `sum(n)`))%>%
  mutate(n=n+`sum(n)`)%>%
  select(-`sum(n)`)
final
```

```{r}
ggplot(data = final, mapping = aes(x=lang,y=n))+
  geom_bar(stat = "identity",fill="blue", color = "black")+
  ggtitle("User ID Distribution")+
  xlab("User ID")+
  ylab("Count")
```
*Observations*

* The majority of accounts are in English, of the rest, half have no language set and half use another language 



### Location

```{r}
top_30 = bots_cleaned%>%
  count(location)%>%
  slice_max(n,n=31)
top_30
# 13040 unkown accounts
sum(top_30$n)-13040

top_100 = bots_cleaned%>%
  count(location)%>%
  slice_max(n,n=101)
sum(top_100$n)-13040

top_1000 = bots_cleaned%>%
  count(location)%>%
  slice_max(n,n=1001)
sum(top_1000$n)-13040

a = bots_cleaned%>%
  count(location)
nrow(a)
```

*Observations*

* Around one third of the accounts have an unknown location (13040)
* Top 30 locations (30/12030=0.2%) have 4003 accounts (4003/37438=10.7%)
* Top 100 locations (100/12030=0.8%) have 6681 accounts (6681/37438=17.8%)
* Top 1000 locations (1000/12030=8.3%) have 12698 accounts (12698/37438=33.9%)



### Username
Username length
```{r}
ggplot(data = bots_cleaned, mapping = aes(str_length(screen_name)))+
  geom_histogram(fill="darkblue", color = "black", binwidth = 1)+
  ggtitle("Username Length Histogram")+
  xlab("Username Length")+
  ylab("Count")+
  theme_minimal()

ggplot(data = bots_cleaned, mapping = aes(str_length(screen_name)))+
  geom_boxplot(fill="darkblue", color = "black")+
  ggtitle("Username Length Box Plot")+
  xlab("Username Length")+
  theme(panel.background = element_rect(fill = "white"))+
  theme_minimal()

summary(str_length(bots_cleaned$screen_name))
```

Digits in Username
```{r}
ggplot(data = bots_cleaned, mapping = aes(str_count(screen_name,"\\d")))+
  geom_histogram(fill="darkblue", color = "black", binwidth = 1)+
  ggtitle("Username Digit Count")+
  xlab("Number of Digits in Username")+
  ylab("Count")+
  theme_minimal()
```
Digits in Username without zero
```{r}
ggplot(data = bots_cleaned%>%filter(str_count(screen_name,"\\d")!=0), mapping = aes(str_count(screen_name,"\\d")))+
  geom_histogram(fill="darkblue", color = "black", binwidth = 1)+
  ggtitle("Username Digit Count")+
  xlab("Number of Digits in Username")+
  ylab("Count")+
  theme_minimal()
```


*Observations*

* Username length is strongly skewed to the left, with a maximum at 15 characters
* The majority of Usernames contain zero digits
* Among usernames with digits, there is a right skew, with a peak at 2 digits



### Verified

```{r}
ggplot(data = bots_cleaned, mapping = aes(verified))+
  geom_histogram(stat= "count", fill="darkblue", color = "black")+
  ggtitle("Verified Distribution")+
  xlab("Verified")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Under one quarter of the Twitter accounts are verified


### Account Age

```{r}
ggplot(data = bots_cleaned, mapping = aes(account_age_days))+
  geom_histogram(fill="blue", color = "black", binwidth = 100)+
  ggtitle("Account Age Distribution")+
  xlab("Account Age")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Account age is skewed left with a peak around 4100 days


### Bot Label

```{r}
ggplot(data = bots_cleaned, mapping = aes(is_bot))+
  geom_bar(fill="blue", color = "black")+
  ggtitle("Bot Label Distribution")+
  xlab("Is Bot?")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Around one third of the Twitter accounts in this data set are bots


### Split

```{r}
ggplot(data = bots_cleaned, mapping = aes(split))+
  geom_bar(fill="darkblue", color = "black")+
  ggtitle("Split Distribution")+
  xlab("Split")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* Almost all split values are 0, with the rest being equally distributed between 1 and 2


### Favourites Count

```{r}
ggplot(data = bots_cleaned, mapping = aes(log_favourites_count))+
  geom_histogram(fill="darkblue", color = "black", binwidth = .1)+
  ggtitle("Favourites Count Distribution")+
  xlab("Favourites Count")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* The transformed and winsorized favorites count is skewed left with a sharp peak around 3.2


### Followers Count

```{r}
ggplot(data = bots_cleaned, mapping = aes(log_followers_count))+
  geom_histogram(fill="darkblue", color = "black", binwidth = .1)+
  ggtitle("Followers Count Distribution")+
  xlab("Followers Count")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* The transformed and winsorized followers count is non-normal and skewed right


### Friends Count

```{r}
ggplot(data = bots_cleaned, mapping = aes(log_friends_count))+
  geom_histogram(fill="darkblue", color = "black", binwidth = .1)+
  ggtitle("Friends Count Distribution")+
  xlab("Friends Count")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* The transformed and winsorized friends count is non-normal and skewed left


### Statuses Count

```{r}
ggplot(data = bots_cleaned, mapping = aes(log_statuses_count))+
  geom_histogram(fill="darkblue", color = "black", binwidth = .1)+
  ggtitle("Statuses Count Distribution")+
  xlab("Statuses Count")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* The transformed and winsorized statuses count is non-normal and skewed left


### Average Tweets Per Day

```{r}
ggplot(data = bots_cleaned, mapping = aes(log_avg_tweets_per_day))+
  geom_histogram(fill="darkblue", color = "black", binwidth = .1)+
  ggtitle("Average Tweets Per Day Distribution")+
  xlab("Average Tweets Per Day")+
  ylab("Count")+
  theme_minimal()
```
*Observations*

* The transformed and winsorized Average Tweets Per Day count is non-normal and skewed right



## Features vs Target

This section performs a bi-variate analysis of the distribution of Bot accounts. This can help in determining what data and feature selection to perform.




### Geography Enabled
```{r}
ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=geo_enabled, fill = is_bot))+
  geom_bar()+
  theme_minimal()
```

```{r}
summary_table <- bots_cleaned %>%
  group_by(geo_enabled, is_bot) %>%
  summarize(count = n()) 

summary_table

test_result <- prop.test(
  x = c(2716, 9709), 
  n = c(2716+14361, 9709+10652)
)
test_result
```


*Observations*

* There is a significantly higher proportion of bots among the accounts with geography disabled (p-value < 2.2e-16)

*Impact*

* geo_enabled is a strong predictor of whether the account is a bot




### Verified
```{r}
ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=verified, fill = is_bot))+
  geom_bar()+
  theme_minimal()
```

```{r}
summary_table2 <- bots_cleaned %>%
  group_by(verified, is_bot) %>%
  summarize(count = n()) 

summary_table2

test_result2 <- prop.test(
  x = c(12109, 316), 
  n = c(17778+12109, 7235+316)
)
test_result2
```
*Observations*

* There is a significantly higher proportion of bots among the accounts that are not verified (p-value < 2.2e-16)

*Impact*

* verified is a significant predictor of whether the account is a bot


### Default Profile
```{r}
ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=default_profile, fill = is_bot))+
  geom_bar()+
  theme_minimal()
```

```{r}
summary_table3 <- bots_cleaned %>%
  group_by(default_profile, is_bot) %>%
  summarize(count = n()) 

summary_table3

test_result3 <- prop.test(
  x = c(7741, 4684), 
  n = c(7979+7741, 17034+4684)
)
test_result3
```
*Observations*

* There is a significantly higher proportion of bots among the accounts that use the default profile (p-value < 2.2e-16)

*Impact*

* Default profile is a significant predictor of whether the account is a bot

### Default Profile Image
```{r}
ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=default_profile_image, fill = is_bot))+
  geom_bar()+
  theme_minimal()
```

```{r}
summary_table4 <- bots_cleaned %>%
  group_by(default_profile_image, is_bot) %>%
  summarize(count = n()) 

summary_table4

test_result4 <- prop.test(
  x = c(383, 12042), 
  n = c(383+175, 12042+24838)
)
test_result4
```
*Observations*

* There is a significantly higher proportion of bots among the accounts that are use the default profile image (p-value < 2.2e-16)

*Impact*

* Default profile image is a significant predictor of whether the account is a bot


### Split
```{r}
ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=split, fill = is_bot))+
  geom_bar()+
  theme_minimal()
```

```{r}
summary_table5 <- bots_cleaned %>%
  group_by(split, is_bot) %>%
  summarize(count = n()) 

summary_table5

# split 0 vs split 1
test_result5a <- prop.test(
  x = c(9932, 1267), 
  n = c(9932+20018, 1267+2476)
)
test_result5a

# split 1 vs split 2
test_result5b <- prop.test(
  x = c(1226, 1267), 
  n = c(1226+2519, 1267+2476)
)
test_result5b

# split 0 vs split 2
test_result5c <- prop.test(
  x = c(1226, 9932), 
  n = c(1226+2519, 9932+20018)
)
test_result5c


```
*Observations*

* There is a no significant difference of proportion of bots among the accounts based on split value (p-value = 0.6153, 0.3187, and 0.41)

*Impact*

* Data set can be filtered out to exclude split because it is not a significant predictor of whether the account is a bot


## Hypothesis

### Are younger accounts with more tweets per day more likely to be bots?

```{r}

ggplot(data = bots_cleaned%>%mutate(is_bot=as.logical(is_bot)), mapping = aes(x=cut(account_age_days, breaks = seq(0, max(account_age_days), by = 500), include.lowest = TRUE), y=log_avg_tweets_per_day, fill = is_bot))+
  geom_boxplot()+
  theme_minimal() +
  labs(x = "Account Age (500-day intervals)", y = "Log Avg Tweets per Day", fill = "Is Bot") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
thing = bots_cleaned%>%
  filter(account_age_days<=500)
thing$is_bot
```
*Observations*

* The very oldest accounts do post more per day than the more recent ones 
* Overall, bot accounts appear to post less than real accounts
* All (three) bots that are less than 500 days old are bots

*Insights*

* Filter data set to remove bots less than 500 days old
* Younger bots with more tweets per day are not more likely to be bots


### Are account activity data significant predictors of bots?
```{r}
activity_model = glm(is_bot ~ log_favourites_count+log_followers_count+log_friends_count+log_statuses_count+log_avg_tweets_per_day, data = bots_cleaned, family = binomial)

summary(activity_model)
```
*Observations*

* Each of the predictors has a significant P value
* Favourites, Followers, and Friends are negativly correlated
* Statuses and Daily Tweets are positively correlated

*Insights*

* Account activity data are significant predictors of bots


## Summary

### Main Insights/Observations

* The year created distribution is skewed right, with a peak in 2009.
* Somewhat less than half of the accounts have default profiles
* Very few accounts have the default profile image
* There is a somewhat bimodal distribution of description length, with peaks around 25 and 160 characters
* A slight majority of accounts do not have geography enabled
* Id values are strongly skewed right
* The majority of accounts are in English, of the rest, half have no language set and half use another language
* Top 100 locations (100/12030=0.8%) have 6681 accounts (6681/37438=17.8%)
* Username length is strongly skewed to the left, with a maximum at 15 characters
* Under one quarter of the Twitter accounts are verified
* The transformed and winsorized favorites count is skewed left with a sharp peak around 3.2
* The transformed and winsorized followers count is non-normal and skewed right
* The transformed and winsorized friends count is non-normal and skewed left
* The transformed and winsorized statuses count is non-normal and skewed left
* The transformed and winsorized Average Tweets Per Day count is non-normal and skewed right
* There is a significantly higher proportion of bots among the accounts with geography disabled (p-value < 2.2e-16)
* There is a significantly higher proportion of bots among the accounts that are not verified (p-value < 2.2e-16)
* There is a significantly higher proportion of bots among the accounts that use the default profile (p-value < 2.2e-16)
* There is a significantly higher proportion of bots among the accounts that are use the default profile image (p-value < 2.2e-16)
* There is a no significant difference of proportion of bots among the accounts based on split value (p-value = 0.6153, 0.3187, and 0.41)
* Younger bots with more tweets per day are not more likely to be bots
* Account activity data are significant predictors of bots


### Implications for Modelling
* geo_enabled is a strong predictor of whether the account is a bot
* verified is a significant predictor of whether the account is a bot
* Default profile is a significant predictor of whether the account is a bot
* Default profile image is a significant predictor of whether the account is a bot
* Data set can be filtered out to exclude split because it is not a significant predictor of whether the account is a bot
* Filter data set to remove bots less than 500 days old




