---
title: "Feature Engineering"
author: "Michal"
date: "2024-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
```

## Data

```{r}
source("cleaning_function.R")
```

```{r}
uncleaned_bots = read.csv("C:/Users/Owner/Desktop/Data Science Project/Twitter-Bot-Code/archive/twitter_human_bots_dataset.csv")
```

```{r}
bots_cleaned = cleaning_function(uncleaned_bots)
# applying EDA recommendations
bots_FE = bots_cleaned%>%
  select(-split)%>%
  filter(account_age_days>500)%>%
  mutate(is_bot=as.logical(is_bot))
```

## Features


### Description Features

Has description - it seems likely that mass producers of bot accounts would not have time to write a description

Description length - perhaps bots use a similar format of descriptions, which would mean that certain lengths have more bots

Number of spaces in description - is it possible that bot account descriptions have a similar amount of words to each other, if they are created based on a template

Periods in Description - the bot account description templates could have a constant number of periods (sentences)

Punctuation in Description - the bot account description templates could have a constant number of punctuation

Bot in Description - the account might admit to being a bot in the description
```{r}
with_features = bots_FE%>%
  mutate(has_desc=str_length(description)!=0)%>%
  mutate(desc_length=str_length(description))%>%
  mutate(desc_space_count=str_count(description,"\\s"))%>%
  mutate(desc_period_count=str_count(description,"\\."))%>%
  mutate(desc_punct_count=str_count(description,"[:punct:]"))%>%
  mutate(bot_in_desc=str_detect(`description`,"bot"))
```
### Name Features
Four features seem reasonable from the Name category. Username length might be a predictor of is_bot, since bots may be named according to certain formats. Additionally, number of digits and punctuation may also predict if the account is a bot, for the same reason. Finally, a lot of bots may admit to being bots in their name, so a marker for the word "bot" should also be added.
```{r}
with_features = with_features%>%
  mutate(username_length=str_length(screen_name))%>%
  mutate(user_digit_count=str_count(screen_name,"\\d"))%>%
  mutate(user_punct_count=str_count(description,"[:punct:]"))%>%
  mutate(bot_in_name=str_detect(`screen_name`,"bot"))
```
### Friends to Followers Ratio
Bots often exist to drive engagement with real accounts (by following them), but do not provide any value, and therefore have few if any followers.
```{r}
with_features = with_features%>%
  mutate(friends_followers_ratio=log_friends_count/log_followers_count)
```
### Location Feature
It makes sense that bots would have their location set to unknown.
```{r}
with_features = with_features%>%
  mutate(unknown_loc=str_detect(location,"unknown"))
```
### Language Feature
It would make sense that bots would not have a language set.
```{r}
with_features = with_features%>%
  mutate(missing_lang = lang == "")
head(with_features)
```


## Feature Selection 
Feature selection is done in two steps:

* Quality Checks - does the feature have a lot of missing values (>95%)
* Correlation analysis - if features are highly correlated with one another (>95%), choose feature with greatest variance

First, creating new data frame with just newly engineered features:
```{r}
only_new_features=with_features%>%
  select(desc_length,desc_space_count,desc_period_count,desc_punct_count,bot_in_desc,username_length,user_digit_count,bot_in_name,friends_followers_ratio,unknown_loc,missing_lang,is_bot)

colnames(only_new_features)
```

### Quality 
```{r}
missing_rows_df=only_new_features%>%
  mutate(missing_rows = rowSums(is.na(.)) > 0)%>%
  filter(missing_rows==TRUE)
nrow(missing_rows_df)
```
No missing values: no need to remove features for quality issues

### Correlation

```{r}
correlation = cor(only_new_features)
corrplot(correlation, method = 'color', type = "lower")
```
Only several of the features are strongly correlated. Make a dataframe with only the correlated features.
```{r}
cor_feature = only_new_features%>%
  select(-bot_in_desc,-username_length,-user_digit_count,-bot_in_name,-friends_followers_ratio)

correlation2 = cor(cor_feature)
corrplot(correlation2, method = 'number', type = "lower")
```
Description length and Description Space Count are most highly correlated (0.90). However, even this is below the 0.95 benchmark, so no features have to be removed.




## EDA for Remaining Engineered Features
Has Description
```{r}
ggplot(data = bots_FE, mapping = aes(x=str_length(description)!=0,fill=is_bot))+
  geom_histogram(stat="count",color = "black")+
  ggtitle("Accounts with Description")+
  xlab("Has a Description")+
  ylab("Count")+
  theme_minimal()


has_description_model = glm(is_bot~has_desc, data = with_features, family = binomial)

summary(has_description_model)
```

Description length
```{r}
ggplot(data = bots_FE, mapping = aes(x=str_length(description),fill=is_bot))+
  geom_histogram()+
  ggtitle("Description Length Histogram")+
  xlab("Description Length")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~str_length(description),data=bots_FE,family = binomial)
summary(model)
```

Number of spaces in description
```{r}
ggplot(data = bots_FE%>%filter(description!=""), mapping = aes(str_count(description,"\\s"),fill=is_bot))+
  geom_histogram()+
  ggtitle("Space Count Histogram")+
  xlab("Space Count")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~str_count(description,"\\s"),data=bots_FE,family = binomial)
summary(model)
```

Periods in Description
```{r}
ggplot(data = bots_FE%>%filter(description!=""), mapping = aes(str_count(description,"\\."),fill=is_bot))+
  geom_histogram()+
  ggtitle("Description Period Count")+
  xlab("Number of Periods in Description")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~str_count(description,"\\."),data=bots_FE,family = binomial)
summary(model)
```

Punctuation in Description
```{r}
ggplot(data = bots_FE%>%filter(description!=""), mapping = aes(str_count(description,"[:punct:]"), fill=is_bot))+
  geom_histogram()+
  ggtitle("Description Punctuation Count")+
  xlab("Number of Punctuation in Description")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~str_count(description,"[:punct:]"),data=bots_FE,family = binomial)
summary(model)
```

Bot in Description
```{r}
ggplot(data = with_features, mapping = aes(bot_in_desc, fill=is_bot))+
  geom_bar()+
  ggtitle("Bot in description vs being bot")+
  xlab("Word bot in Description")+
  ylab("Count")+
  scale_y_log10()+
  theme_minimal()

model = glm(is_bot~bot_in_desc,data=with_features,family = binomial)
summary(model)
```

Username Length - not significant, to be removed
```{r}
ggplot(data = with_features, mapping = aes(username_length, fill=is_bot))+
  geom_histogram()+
  ggtitle("Username Length")+
  xlab("Number of Characters in Username")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~username_length,data=with_features,family = binomial)
summary(model)
```

Username Digit Count
```{r}
ggplot(data = with_features, mapping = aes(user_digit_count, fill=is_bot))+
  geom_histogram()+
  ggtitle("Username Digits")+
  xlab("Number of Digits in Username")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~user_digit_count,data=with_features,family = binomial)
summary(model)
```

Username Punctuation
```{r}
ggplot(data = with_features, mapping = aes(user_punct_count, fill=is_bot))+
  geom_histogram()+
  ggtitle("Username Punctuation")+
  xlab("Number of Punctuation in Username")+
  ylab("Count")+
  theme_minimal()

model = glm(is_bot~user_punct_count,data=with_features,family = binomial)
summary(model)
```

Bot in Username
```{r}
ggplot(data = with_features, mapping = aes(bot_in_name, fill=is_bot))+
  geom_bar()+
  ggtitle("Bot in name vs being bot")+
  xlab("Word bot in Username")+
  ylab("Count")+
  scale_y_log10()+
  theme_minimal()

model = glm(is_bot~bot_in_name,data=with_features,family = binomial)
summary(model)
```

Friends to Followers Ratio
```{r}
ggplot(data = with_features, mapping = aes(friends_followers_ratio, fill=is_bot))+
  geom_histogram()+
  ggtitle("Friends to Followers Ratio")+
  xlab("Ratio")+
  ylab("Count")+
  scale_y_log10()+
  theme_minimal()

model = glm(is_bot~friends_followers_ratio,data=with_features,family = binomial)
summary(model)
```

Location Unknown
```{r}
ggplot(data = with_features, mapping = aes(unknown_loc, fill=is_bot))+
  geom_bar()+
  ggtitle("Unkown Location")+
  xlab("Unkown Location")+
  ylab("Count")
  theme_minimal()

model = glm(is_bot~unknown_loc,data=with_features,family = binomial)
summary(model)
```

Missing Language
```{r}
ggplot(data = with_features, mapping = aes(missing_lang, fill=is_bot))+
  geom_bar()+
  ggtitle("Missing Language")+
  xlab("Missing Language")+
  ylab("Count")
  theme_minimal()

model = glm(is_bot~missing_lang,data=with_features,family = binomial)
summary(model)
```


Observations
* Most of the engineered features are significantly correlated to the probability of the account being a bot

Impact
* Username Length is not significant so it can be dropped

```{r}
final_df = with_features%>%
  select(-username_length)
colnames(final_df)
```


## Feature Engineering PipeLine
```{r}
feature_function = function(data){
  # run data cleaning function
  cleaned = data%>%
   #  remove unnecessary columns
    select(-X,-profile_image_path,-profile_background_image_path, -profile_background_image_url, -profile_image_url)%>%
   
     # standardize missing location  value
    mutate(
      location = if_else(str_trim(location) == "","unknown",location))%>%
  
    # type cast to correct class
  mutate(
    created_at = ymd_hms(created_at),
    default_profile = as.logical(default_profile),
    default_profile_image = as.logical(default_profile_image),
    geo_enabled = as.logical(geo_enabled),
    verified = as.logical(verified))%>%
    
     #cast and rename account_type
  mutate(
    account_type=if_else(account_type=="bot",1,0))%>%
  rename(is_bot=account_type)%>%
    
    # transform and winsorize outliers
    mutate(
      # log(x+1) to avoid "-inf"
      log_favourites_count = log10(favourites_count+1),
      log_followers_count = log10(followers_count+1),
      log_friends_count = log10(friends_count+1),
      log_statuses_count = log10(statuses_count+1),
      log_avg_tweets_per_day = log10(average_tweets_per_day+1)
    )%>%
    mutate(
      log_favourites_count= Winsorize(log_favourites_count),
      log_followers_count = Winsorize(log_followers_count),
      log_friends_count = Winsorize(log_friends_count),
      log_statuses_count = Winsorize(log_statuses_count),
      log_avg_tweets_per_day = Winsorize(log_avg_tweets_per_day)
    )%>%
  
  # remove old columns
  select(-favourites_count,-followers_count,-friends_count, -statuses_count, -average_tweets_per_day)
  
  
  
  final = cleaned %>% 
    # apply EDA recommendations
    select(-split)%>%
    filter(account_age_days>500)%>%
    # add new features
    mutate(is_bot=as.logical(is_bot))%>%
    mutate(has_desc=str_length(description)!=0)%>%
    mutate(desc_length=str_length(description))%>%
    mutate(desc_space_count=str_count(description,"\\s"))%>%
    mutate(desc_period_count=str_count(description,"\\."))%>%
    mutate(desc_punct_count=str_count(description,"[:punct:]"))%>%
    mutate(bot_in_desc=str_detect(`description`,"bot"))%>%
    mutate(user_digit_count=str_count(screen_name,"\\d"))%>%
    mutate(user_punct_count=str_count(description,"[:punct:]"))%>%
    mutate(bot_in_name=str_detect(`screen_name`,"bot"))%>%
    mutate(friends_followers_ratio=log_friends_count/log_followers_count)%>%
    mutate(unknown_loc=str_detect(location,"unknown"))%>%
    mutate(missing_lang = lang == "")
  
  return(final)
}
```

## Summary

#### Feature Engineering Summary

* 14 new features were engineered, mostly relating to the string variables
* one mathematical ratio (friends to followers) was added
* one feature was dropped due to lack of significant correlation


#### Implications of Modeling

* Engineered and selected 13 features that should theoretically be useful in the prediction model
* Feature engineering pipeline was designed, so new data can be easily transformed
